
---
title: "SEM Data Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document provides an example of how to analyze a Structural Equation Modeling (SEM) dataset. The dataset contains observed indicators for two latent variables (*Cognitive Load* and *Performance*) and additional observed variables (*Stress*, *Age*, and *Gender*).

---

## Step 1: Load the Data

```{r load-data}
# Load necessary libraries
library(tidyverse)

# Load the SEM dataset (assuming it's in the same folder as this Rmd file)
data <- read.csv("sem_data.csv")

# Quick look at the data
head(data)
summary(data)
```

---

## Step 2: Explore and Visualize the Data

```{r explore-data}
# Scatterplot of Cognitive Load indicators
ggplot(data, aes(x = Cognitive_Load_1, y = Cognitive_Load_2)) +
  geom_point(color = "blue") +
  labs(title = "Cognitive Load Indicators", x = "Cognitive_Load_1", y = "Cognitive_Load_2")

# Scatterplot of Performance indicators
ggplot(data, aes(x = Performance_1, y = Performance_2)) +
  geom_point(color = "green") +
  labs(title = "Performance Indicators", x = "Performance_1", y = "Performance_2")

# Boxplot of Stress by Gender
ggplot(data, aes(x = Gender, y = Stress)) +
  geom_boxplot(fill = "orange", color = "black") +
  labs(title = "Stress Levels by Gender", x = "Gender", y = "Stress")
```

---

## Step 3: Prepare Data for SEM

```{r prepare-data}
# Select only the observed indicators for SEM
sem_data <- data %>%
  select(Cognitive_Load_1, Cognitive_Load_2, Cognitive_Load_3, Performance_1, Performance_2, Performance_3)

# Check the structure of the SEM data
summary(sem_data)
```

---

## Step 4: Perform SEM Analysis

```{r sem-analysis, message=FALSE, warning=FALSE}
# Load SEM library
library(lavaan)

# Define the SEM model
sem_model <- '
  # Latent variables
  Cognitive_Load =~ Cognitive_Load_1 + Cognitive_Load_2 + Cognitive_Load_3
  Performance =~ Performance_1 + Performance_2 + Performance_3
'

# Fit the SEM model
sem_fit <- sem(sem_model, data = sem_data)

# Summary of the SEM model
summary(sem_fit, fit.measures = TRUE, standardized = TRUE)
```

---

## Step 5: Interpret Results

```{r interpret-results}
# Display fit measures
fit_measures <- fitMeasures(sem_fit)
print(fit_measures)

# Extract standardized estimates
std_estimates <- standardizedSolution(sem_fit)
print(std_estimates)
```

---

## Conclusion

This document provides a basic example of SEM analysis using R. You can expand it by adding more latent variables, testing alternative models, or visualizing the results. Feel free to adapt this workflow to your own research!
